---
title: "Bateries"
output: html_notebook
---
Research question: Is there a difference in the battery lifetime for different types of material and different operating temperatures?
Solution: We will use a two-way anova with replication - how does two factors (temperature and type of material) effect battery lifetime.

```{r}
install.packages("ggpubr") #this one is needed for the interaction plot
install.packages("tidyverse")
```
We need to import our data for analysis. 
```{r}
library(readxl)
baterie <- read_excel("baterieVZ.xlsx")
View(baterie)
```



Next we want to see the internal structure of our data.
```{r}
str(baterie)
```
For the useage of anova we need to convert "type of material" and "temperature" as factors.
```{r}
baterie$material <- factor(baterie$material)
baterie$teplota <- factor(baterie$teplota)
                  
head(baterie)
str(baterie)
```


## Visualization

# Box plots

          
```{r}
library("ggpubr")
ggboxplot(baterie, x = "material", y = "zivotnost", color = c("#999999", "#E69F00", "#56B4E9"), main = "Dependency of battery lifetime and type of material", xlab = "Type of material", ylab = "Battery lifetime") +
  geom_jitter(width = 0.1)
```

```{r}
library("ggpubr")
ggboxplot(baterie, x = "teplota", y = "zivotnost", color = c("#999999", "#E69F00", "#56B4E9"), main = "Dependency of battery lifetime and temperature", xlab = "Temperature", ylab = "Battery lifetime") +
  geom_jitter(width = 0.1)
```


Combined into one box plot: lifetime from type of material and lifetime from temperature 
```{r}
boxplot(zivotnost ~ material * teplota, data = baterie, frame = FALSE, col = c("white", "steelblue"), xlab = "Type of material.Temperature", ylab = "Battery lifetime")
```

# Interaction plots
Interaction plot is suitable for representing relatonships between lifetime of bateries, types of material and temperature.
For creating an interaction plot we can use a ggline or a interaction.plot function.
       
```{r}
library("ggpubr")
ggline(baterie, x = "teplota", y = "zivotnost", color = "material", add = c("mean_se"), xlab = "Temperature", ylab = "Battery lifetime", palette = c("#999999", "#E69F00", "#56B4E9"), main = "Interaction plot (with standard error)")
```


                 
```{r}
interaction.plot(x.factor = baterie$teplota, trace.factor = baterie$material, response = baterie$zivotnost, fun = mean, type = "b", legend = TRUE, xlab = "Temperature", ylab = "Battery lifetime", col = c("#999999", "#E69F00", "#56B4E9"), main = "Interaction plot")
```

# Two-way anova with interaction effect

```{r}
anova <- aov(zivotnost ~ material * teplota, data = baterie) # Two-way ANOVA with interaction effect
summary(anova)
```
From the results of anova, we can conclude the following (based on the p-values and a significance level of 0.05):
The p-value of type of material (material) and temperature (teplota) is significant. Also the interaction of both factors is significant, from which we conclude that these factors influence each other.



